---
title: "Sampling Distribution"
author: "Aanish Singla"
date: "1 April 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Overview**: In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution is simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials. Note that we will do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. 

##Part 1:
Now look at the Theoretical mean. Mean = 1/lambda, which in this case is 1/.2 = 5

Now lets generate the distribution of means of 1000 samples of exp distribution of size 40 and lambda - 0.2

```{r}
sample_means = NULL
  for (i in 1 : 1000) 
    sample_means = c(sample_means, mean(rexp(40,0.2)))
mean(sample_means)
```

Lets see this visually

```{r fig.width=4, fig.height=3}
hist(sample_means, breaks=100)
  abline(v=mean(sample_means), lwd=2)
```

**Pretty close to the theoretical value of the mean!**

##Part 2:
Now lets simulate the variance of the means. Theoretical Standard error of the mean = sigma/sqrt(n), which is 5/sqrt(1000)

```{r}
   5/sqrt(1000)
```

Since variance is square of standard error of mean, Variance is
```{r}
   (5/sqrt(1000))^2
```

Now lets calculate the variance of the sample that we have:</p>

```{r}
  Var <- (sd(sample_means)/sqrt(1000))   
  Var  
```
**Pretty close to the theoretical value of the variance!**

##Part 3:
Lets see if our results comply with central limit theorem

Distribution of large collection (1000) of random exponentials looks as follows
```{r fig.width=4, fig.height=3}
   hist(rexp(1000,0.2), breaks=100)
```
  
Now lets plot the distribution of sample means of 40 exponentials
```{r fig.width=4, results='hide',fig.width=4, fig.height=3}
  myhist <- hist(sample_means, breaks=100)
```

Overlaying a normal curve:
```{r fig.width=4, fig.height=3}
  multiplier <- myhist$counts / myhist$density
  mydensity <- density(sample_means)
  mydensity$y <- mydensity$y * multiplier[1]
  
  plot(myhist)
  lines(mydensity, lwd=2)
```
    
**Viola! central limit theorem works :-)**
