---
title: "Tooth Growth Data Analysis"
author: "Aanish Singla"
date: "1 April 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Overview**: In this document we will analyze the ToothGrowth dataset that comes with R "datasets" and do hypothesis testing for facts which are not obvious.

Lets see variables and the values:
```{r}
library("datasets")
str(ToothGrowth)
```

There are 3 variables - len, supp and dose. Supp (Delivery Method) is a 2 level factor with value "VC" and "OJ". Other 2 fields are numeric.

Also we can see that there are 60 rows in the dataset.

Lets see how the plot looks for length and dose for different supp

```{r fig.width=4, fig.height=3}
library(ggplot2)
g <- ggplot(ToothGrowth, aes(dose, len))
g + geom_point(aes(color=supp), size=2) + 
  xlab("Dose") + ylab("Length") + ggtitle("Tooth Growth Data")
```

There is a clear increase in length as the Dose increases.   


**What about the effect of supp(delivery method) on length?**

Lets frame a hypothesis for this. 
Null Hypothesis: H0 = Delivery method has no effect on length i.e.
                 H0 = Mean length with OJ = Mean  length with VC
                 
Alternate Hypothesis = Ha = Delivery method has an effect on length i.e.                   Ha = Mean length with OJ <> Mean  length with VC

Lets perform a two sided non-paired T-test to verify the Hypothesis (assuming variances are not equal)    

```{r}
t.test(ToothGrowth$len~ToothGrowth$supp, paired=F, 
       var.equal=F, alt="two.sided")
```

**Since the p-value is 0.06063 (greater than 0.05), we fail to reject the Null hypothesis. Also we can notice that that 95% confidence interval contains zero in it, so we have no statistical evidence that one delivery method is better than other.**